---
title: "Untitled"
author: "Manaswini"
date: "2024-04-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


#Required libraries
```{r}

library(cluster)
library(caret)
library(dendextend)
library(knitr)
library(factoextra)
library(readr)

```


#Importing the dataset
```{r}

df <- read.csv("C:/Users/Winnie/Downloads/Cereals.csv")
View(df)
data <- data.frame(df[,4:16])

```


#Removing missing values of all cereals
```{r}
data <- na.omit(data)
```


#Data normalization using scale function
```{r}
data_normalize <- scale(data)
```


##TASK-1
#Using Euclidean distance to create hierarchial cluster using normalized data
#Complete Linkage with hierarchial cluster
```{r}
Dist <- dist(data_normalize, method = "euclidean")

H_clust <- hclust(Dist, method = "complete")
```


#Dendogram Plot
```{r}

plot(H_clust, cex = 0.7, hang = -1)

```


#Perform computations using AGNES clustering algorithm and with various linkage methods
```{r}
single_Hclust <- agnes(data_normalize, method = "single")
complete_Hclust <- agnes(data_normalize, method = "complete")
average_Hclust <- agnes(data_normalize, method = "average")
ward_Hclust <- agnes(data_normalize, method = "ward")
```

#Choosing the correct course of action
```{r}

print(single_Hclust$ac)
print(complete_Hclust$ac)
print(average_Hclust$ac)
print(ward_Hclust$ac)

```
#ward startegy is most suitable one, as it's value is 0.9046042


#TASK-2
#choosing the number of clusters
```{r}

pltree(ward_Hclust, cex = 0.5, hang = -1, main = "Dendrogram of agnes (Using Ward)")
rect.hclust(ward_Hclust, k = 5, border = 2:7)
R_Group <- cutree(ward_Hclust, k=5)
D_frame_2 <- as.data.frame(cbind(data_normalize,R_Group))

```



```{r}
fviz_cluster(list(data = D_frame_2, cluster = R_Group))
```
#From above observation 5 clusters can be choosen



#TASK-3
#Determining the stability and structure of the clusters. 
```{r}
#creating partitions
set.seed(123)
partition_one <- data[1:55,]
partition_two <- data[56:74,]
```


#Performing Hierarchical Clustering while considering k = 5. Compute with AGNES and with different linkage methods for training dataset
```{r}

single_df <- agnes(scale(partition_one), method = "single")
complete_df <- agnes(scale(partition_one), method = "complete")
average_df <- agnes(scale(partition_one), method = "average")
ward_df <- agnes(scale(partition_one), method = "ward")
cbind(single=single_df$ac , complete=complete_df$ac , average= average_df$ac , ward= ward_df$ac)
pltree(ward_df, cex = 0.6, hang = -1, main = "Dendogram of Agnes with Partitioned Data (Using Ward)")
rect.hclust(ward_df, k = 5, border = 2:7)
cut_2 <- cutree(ward_df, k = 5)

```


#Calculating the centroids
```{r}

data_result <- as.data.frame(cbind(partition_one, cut_2))
data_result[data_result$cut_2==1,]
centroid_1 <- colMeans(data_result[data_result$cut_2==1,])
data_result[data_result$cut_2==2,]
centroid_2 <- colMeans(data_result[data_result$cut_2==2,])
data_result[data_result$cut_2==3,]
centroid_3 <- colMeans(data_result[data_result$cut_2==3,])
data_result[data_result$cut_2==4,]
centroid_4 <- colMeans(data_result[data_result$cut_2==4,])
centroids <- rbind(centroid_1, centroid_2, centroid_3, centroid_4)
x2 <- as.data.frame(rbind(centroids[,-14], partition_two))

```


#figuring out the distance
```{r}

Dist_1 <- get_dist(x2)
Matrix_1 <- as.matrix(Dist_1)
dataframe1 <- data.frame(data=seq(1,nrow(partition_two),1), Clusters = rep(0,nrow(partition_two)))
for(i in 1:nrow(partition_two)) 
  {dataframe1[i,2] <- which.min(Matrix_1[i+4, 1:4])}
dataframe1
cbind(D_frame_2$R_Group[56:74], dataframe1$Clusters)
table(D_frame_2$R_Group[56:74] == dataframe1$Clusters)

```

#From above observation we can conclude that the model is stable as we got 12 TRUE and 7 FALSE


#TASK-4
#The elementary public schools would like to choose a set of cereals to include in their
daily cafeterias. Every day a different cereal is offered, but all cereals should support a
healthy diet. For this goal, you are requested to find a cluster of “healthy cereals.”
Should the data be normalized? If not, how should they be used in the cluster analysis?
```{r}

Health_data <- df
Health_data_RD <- na.omit(Health_data)
clust <- cbind(Health_data_RD, R_Group)
clust[clust$R_Group==1,]
clust[clust$R_Group==2,]
clust[clust$R_Group==3,]
clust[clust$R_Group==4,]

```


```{r}

mean(clust[clust$R_Group==1,"rating"])
mean(clust[clust$R_Group==2,"rating"])
mean(clust[clust$R_Group==3,"rating"])
mean(clust[clust$R_Group==4,"rating"])

```
#From the above observation we can see that cluster 1 has the highest value 
#Hence Group 1 can be considered as the cluster for healthy diet
